---
title: "Data Analysis I"
author: "Zhiyuan Jiang"
date: "28/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gapminder)
library(tidyverse)
library(fs)
```

## Chapter 5 Basic Care and Feeding of Data in R

```{r}

str(gapminder)
class(gapminder)
names(gapminder)
nrow(gapminder)
ncol(gapminder)
dim(gapminder)
length(gapminder)
summary(gapminder)
plot(lifeExp ~ year, gapminder) 
```
Year in the x-axis, against life in y-axis; y ~ x, like any formula 
```{r}
plot(lifeExp ~ gdpPercap, gapminder)
plot(lifeExp ~ log(gdpPercap), gapminder)

hist(gapminder$lifeExp)
summary(gapminder$year)
levels(gapminder$country)
nlevels(gapminder$country)
barplot(table(gapminder$continent))
```

Some basic ggplot2 graph
```{r}
p <- ggplot(filter(gapminder, continent != "Oceania"),
            aes(x = gdpPercap, y = lifeExp))
p <- p + scale_x_log10()
p + geom_point()
p+geom_point(aes(color = continent))
p + geom_point(alpha = (1/3), size = 3) + geom_smooth(lwd = 3, se = FALSE)
p + geom_point(alpha = (1/3), size = 3) + facet_wrap( ~ continent) + 
  geom_smooth(lwd = 1.5, se = FALSE)
```

##Chapter 6 Introduction ot dplyr

```{r}
filter(gapminder, lifeExp < 29)
filter(gapminder, country == "Rwanda", year > 1979)
filter(gapminder, country %in% c("Rwanda", "Afghanistan"))

select(gapminder, year, lifeExp)

gapminder %>% select(year, lifeExp) %>% head(4)
```

##Chapter 7 Singel table dplyr functions
```{r}
gap <- gapminder

gap_can <- gap %>% filter(country == "Canada")

bench <- gap %>% filter(country == "Australia") 

gap <- gap %>% mutate(
  benchmark = rep(bench$gdpPercap, nlevels(country)),
  ratio = gdpPercap/benchmark,
  benchmark = NULL
) 

gap %>% 
  filter(year == 2007) %>%  
  arrange(desc(gdpPercap * pop))

gap %>% 
  rename(life_exp = lifeExp,
         gdp_Percap = gdpPercap)
```

```{r}
gap %>%
  filter(country == "Burundi", year > 1996) %>%
  select(yr = year, lifeExp, gdpPercap) %>%
  select(gdpPercap, everything())
```
we can use select to rename variable, new name is still before the old one
and select will display information in given order, use everything to indicate literally everything.

```{r}
gap %>% 
  group_by(continent) %>% 
  summarise(n = n())

gap %>%
  group_by(continent) %>%
  tally()

gap %>%
  count(continent)

gap %>%
  group_by(continent) %>%
  summarize(n = n_distinct(country))
```

```{r}
gap %>%
  filter(year %in% c(1952, 2007)) %>% 
  group_by(continent, year) %>% 
  summarize_at(vars(lifeExp, gdpPercap), list(~mean(.), ~median(.)))

gap %>%
  filter(continent == "Asia") %>%
  group_by(country) %>%
  summarize(max_life = max(lifeExp), min_life = min(lifeExp))

```
summarize_at() function allowing summarize multiple variables at one shot. The first vars() argument requring different variables, and then indicates the function to use different functions to summarize.

```{r}
gap %>%
  group_by(country) %>%
  select(country, year, lifeExp) %>%
  mutate(lifeExp_diff = lifeExp - first(lifeExp)) %>%
  filter(year == 2007) %>% arrange(lifeExp_diff)
```

The first() function will extract the first value of the vector, like here it extract the first value of lifeExp within each country group.

```{r}
gap %>%
  filter(continent == "Asia") %>% 
  select(year, country, lifeExp) %>% 
  group_by(year) %>% 
  filter(min_rank(desc(lifeExp)) < 2 | min_rank(lifeExp) < 2) %>% 
  arrange(year) %>% 
  print(n = Inf)
  
 asia <- gap %>% filter(continent == "Asia")
 
 asia %>% group_by(year) %>%
   mutate(le_rank = min_rank(lifeExp),
          le_desc_rank = min_rank(desc(lifeExp))) %>% 
   filter(year > 1995)

 gap %>%
   arrange(year) %>%
   group_by(year) %>% 
   top_n(1, wt = desc(lifeExp))
```

```{r}
gap %>% 
  select(country, year, continent, lifeExp) %>% 
  group_by(continent, country) %>% 
  mutate(le_delta = lifeExp - lag(lifeExp)) %>% 
  summarize(worst_le_delta = min(le_delta, na.rm = TRUE)) %>% 
  top_n(-1, wt = worst_le_delta) %>% 
  arrange(worst_le_delta)
```

##Chapter 9 Writing and reading files
```{r}
gap_tsv <- path_package("gapminder", "extdata", "gapminder.tsv")

gapminder <- read_tsv(gap_tsv)

str(gapminder, give.attr = FALSE)


gap_life_exp <- gapminder %>% 
  group_by(country, continent) %>% 
  summarise(life_exp = max(lifeExp)) %>% 
  ungroup()

write_csv(gap_life_exp, "gap_life_exp.csv")

```

to write a data, specify which data frame you want outport, and give it a name, with appropriate suffix

```{r}

gap_life <- gap_life %>% 
  mutate(country = fct_reorder(country, life_exp))

head(levels(gap_life$country))

```